---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.  
2. If you have already settled on a group, please have one member send Bry (brymoore@ucdavis.edu) an email with the names of the group members.
3. If you don't have a group, hang tight we are working on it.  
4. Please setup your computer as normal for today.  

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.  
```{r}
malaria <- read.csv("data/malaria.csv")
```

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.  
```{r}
# Sounds good
```

3. Make the data tidy and store them as a new object.   
```{r}
malaria <- read.csv("data/malaria.csv") %>% 
  clean_names
```


```{r}
malaria_tidy <- malaria %>% 
  pivot_longer(cols = starts_with("malaria_rdt"),
               names_to = "age_class",
               values_to = "cases") %>% 
  select(newid, data_date, submitted_date, location_name, district, age_class, cases)
```


4. Which district had the highest *total* number of cases on July 30, 2020?  
```{r}
malaria_tidy %>% 
  group_by(district) %>% 
  filter(data_date == "2020-07-30") %>% 
  summarise(total_cases = sum(cases, na.rm = T)) %>% 
  arrange(desc(total_cases))
```

#### The Bolo district had the highest total number of cases at 347 on July 30. 2020.
