---
title: "Midterm Cheat Sheet"
author: "Elisabeth Sellinger"
date: "2024-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Lab 1:**  
- intro: set-up  
- lab 1_1: class info  
- lab 1_2: r-markdown info and arithmetic  
- hw1: difference between R and R studio  

**Lab 2:**  
- intro: lab 1 review  
- lab 2_1: Define an object in R, Use objects to perform calculations, Explain the difference between data classes in R, Use R to identify the class of specific data, Define NA in R, Determine whether or not data have NA values.   
- lab 2_2: Define data structures and provide examples of vectors and data matrices, Build a new vector and call elements within it, Combine a series of vectors into a data matrix, Name columns and rows in a data matrix, Select values and use summary functions in a data matrix.    
- hw2: Define vectors and data matrices  

**Lab 3:**  
- intro: Vector and matrix practice  
- lab 3_1: Combine a series of vectors into a data frame, Name columns and rows in a data frame, Select columns and rows and use summary functions, Change data classes of columns in the data frame, Write your data frame to a csv file!   
- lab 3_2: Import .csv files as data frames using `read_csv()`, Use summary functions to explore the dimensions, structure, and contents of a data frame, Use the `select()` command of dplyr to sort data frames.  
- hw3: mammals sleep data into  

**Lab 4:**  
- intro: warm-up with dataframe practice  
- lab 4_1: Use summary functions to assess the structure of a data frame, Use the select function of `dplyr` to build data frames restricted to variable of interest, Use the `rename()` function to provide new, consistent names to variables in data frames.  
- lab 4_2: Use the functions of dplyr (filter, select, arrange) to organize and sort data frames, Use `mutate()` to calculate a new column from existing columns. Between, filtering commands and rules.  
- hw4: homerange data, levels, table  

**Lab 5**  
- intro: select vs filter. warmup lab 4 review  
- lab 5_1: Use `filter()` to extract variables of interest, Use `filter()` to extract variables of interest under multiple conditions.  
- lab 5_2: Use pipes to connect functions in dplyr, Use `arrange()` to order dplyr outputs, Use `mutate()` to add columns in a dataframe, Use `mutate()` and `if_else()` to replace values in a dataframe. 

**Lab 6**  
- intro: warmup: pipe practice  
- lab 6_1: Use `mutate()` to add columns in a dataframe, Use `mutate()` and `if_else()` to replace values in a dataframe.  
- lab 6_2: Develop your dplyr superpowers so you can easily and confidently manipulate dataframes, Learn helpful new functions that are part of the `janitor` package.    
- hw5: same as 6_2  

**Lab 7**  
- intro: lab 6 review  
- lab 7_1: Use a combination of `select()`, `filter()`, and `mutate()` to transform datamframes, Use the `skimr` package to produce summaries of data, Produce clean summaries of data using `summarize()`, Use `group_by()` in combination with `summarize()` to produce grouped summaries of data. Also learn `n_distinct()`.    
- lab 7_2: Produce clear, concise summaries using a variety of functions in `dplyr` and  `janitor.`, Use the `across()` operator to produce summaries across multiple variables. Using `group_by()` for multiple columns. Using `count()`.
- hw6: When to use `group_by()` vs `count()`. Lots of practice

**Lab 8**
- intro: midterm review
- lab 8_1: across review! helpful 'across()' functions
- lab 8_2: Define `NA` and describe how they are treated in R, Produce summaries of the number of NA's in a data set, Replace values with `NA` in a data set as appropriate.  
- hw7:

**Lab 9**
- 


**Quick Code:**  
```{r}
# remove NAs: na.rm = T
# change column class: data$column <- as.factor(data$column)
# select column based if character: select_if(data, is.character)
# rename column: rename(data, newname = "oldname")
# Filtering multiple values within the same variable: filter(data, column %in% c(value, value)) # Between in filtering: filter(data, between(column, value1, value2)
# Filter on multiple conditions: filter(data, column1 == "value" & column2 > value)
# Make all entries lowercase: data %>% mutate_all(tolower)
# Make some values lowercase: data %>% mutate(across(c("column1", "column2"), tolower))
# if_else in filtering: mutate(data = ifelse(column == value1, value2, column)) if equals value1, change it to value2, if not value1, leave it the same
# to see quick stat summary use: skim(data)
# See the total values of a column in summarize(): sum()
# See the distinct number of types in a column: n_distinct
# Determine number of observations in column: count()
# Look at percentages of values in column: tabyl(data$column) -- similar to table()
# 'across()' function: summarize(across(c(column1, column2, column3), function))
```





