---
title: "BIS15L_AI_hmwk"
author: "ELisabeth Sellinger"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunk does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

#### This code chunk loads the necessary libraries for the homework.
```{r}
library(tidyverse)
library(janitor)
library(dplyr)
```
   

1. This code chunk reads in the blood count dataset and had to be edited to include the correct file path name. It also had to be edited to include the "clean_names" function to clean the column names. 
```{r}
blood_count <- read_csv("data/blood_count_dataset.csv") %>% 
  clean_names()
```
 

2. This code chunk had a typo in the function name. It was changed from "glmpse" to "glimpse" to properly display the dataset.
```{r}
glimpse(blood_count)
```
 

3. This code chunk had a typo in the function name. It was changed from "maen" to "mean" to properly display the dataset. It also had a missing comma after the "min" function and a missing closing parenthesis at the end of the code.  
```{r}
blood_count %>% 
  group_by(gender) %>% 
  summarize(min=min(red_blood_cells),
            mean=mean(red_blood_cells),
            max=max(red_blood_cells))
```


4. This code chunk had 2 typos in the "filter" function. It was changed from = to == and `male` to `Male` to properly filter the dataset. It also had a typo in the "select" function. It was changed from [] to () to properly select the columns. I also changed colMeans to summarise_all.
```{r}
options(scipen = 999, digits = 2)

blood_count %>% 
  filter(gender == "Male") %>%
  select(3:9) %>%
  summarise_all(mean, na.rm = TRUE)
```
  

5. This code chunk had a typo in the "filter" function and dataset name. It was changed from = to ==. It also had a typo in the "select" function. It was changed from [] to () to properly select the columns. I also changed colMeans to summarise_all.
```{r}
blood_count %>% 
  filter(gender == "Female") %>%
  select(3:9) %>%
  summarise_all(mean, na.rm = TRUE)
```


6. This code chunk had a typo in the "pivot_wider" function. It was changed from [] to () to properly select the columns. It also had a missing comma after the "facet" function.
```{r dpi = 250, out.width = "75%"}
blood_count %>%
  pivot_longer(cols = (3:9), names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Test, y = Values, fill = gender)) +
  geom_boxplot() + 
  scale_y_log10() + #make y axis on a log scale
  facet_wrap(~ Test, scales = "free") 
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values",
       fill = "Gender")
```


7. This code chunk had a typo in the "mutate" function. It was changed from "change" to "mutate" to properly mutate the dataset. It also had a missing closing parenthesis at the end of the code.
```{r}
grouped_bc <- blood_count %>%
  mutate(Age_Group = case_when(
    age >= 20 & age <= 29 ~ "20s",
    age >= 30 & age <= 39 ~ "30s",
    age >= 40 & age <= 49 ~ "40s",
    age >= 50 & age <= 59 ~ "50s",
    age >= 60 & age <= 70 ~ "60s")) # there were only 4 people in 70s, so I combined it with 60s

head(grouped_bc)
```


8. The code chunk above had a typo in the "ggplot" function and selecting the data. It was changed from "ase" to "aes" to properly plot the data. It also had a typo in the "values" variable. It was changed from "values" to "Values" to properly plot the data. It also had a missing ~ after the "facet" function.
```{r dpi = 250, out.width = "75%"}
grouped_bc %>%
  pivot_longer(cols = 3:9, names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Age_Group, y = Values, fill = Age_Group)) +
  geom_boxplot() +
  facet_wrap(~Test, scales = "free") +  # Facet by the "Test" variable
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values")
```


9. What are two drawbacks and two benefits of AI in coding?
2 drawbacks are that AI is not perfect and doesn't always understand the context of the code. It also can't always understand the intent of the code. 2 benefits are that AI can help catch errors and can help with repetitive tasks.

10. Do you think you will use AI for coding in the future? Why or why not?
I think I will use AI for coding in the future because it can help catch errors and can help with repetitive tasks but I will also be cautious because it definitely doesn't catch everything and doesn't always understand the context of the code.

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 