---
title: "BIS15L_AI_class"
author: "Elisabeth Sellinger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI is the simulation of human intelligence processes by machines, especially computer systems. These processes include learning, reasoning, and self-correction. Specific applications of AI include expert systems, natural language processing (NLP), speech recognition and machine vision.
2. What is ChatGPT?
ChatGPT is a language model that can generate human-like text based on the input it receives. It is a large-scale transformer-based language model that can be used for a variety of natural language processing tasks, such as text generation, translation, summarization, and more.
3. What is Github Copilot?
GitHub Copilot is an AI pair programmer that helps you write code faster. It is powered by OpenAI's GPT-3 and is available as a Visual Studio Code extension. It can help you write code, suggest whole lines or blocks, and even entire functions. It can also help you find and fix bugs, and write tests.
4. What are unicorns?
Unicorns are mythical creatures that are often depicted as horses with a single, spiraling horn on their foreheads. They are often associated with purity and grace, and are a symbol of fantasy and magic. They are also a popular subject in literature, art, and popular culture.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix <- read_csv("Data/netflix_titles.csv")
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*
#### Yes, Copilot helped me write the code above. The prompt was clear and concise.

## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
#use the glimpse function to explore the data
glimpse(netflix)
```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*
#### No I had to use the prompt "use the glimpse function to explore the data" to get the desired output. 

Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <- netflix %>% 
  filter(type == "Movie")
```


Which director has the most movies on netflix?
```{r}
netflix_movies %>% 
  count(director, sort = TRUE)
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*
#### Yes, there are NA's in the director column. 

Which director has the most movies on netflix? Sort out any NAs
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  count(director, sort = TRUE)
```

Which director had the most movies by release year and remove the NA's?
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  count(director, release_year, sort = TRUE)
```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*
#### Yes, the first director has two different names.

Which director had the most movies by release year and remove the NA's? Separate the names by a comma and remove any leading or trailing white spaces.
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  separate_rows(director, sep = ",") %>% 
  mutate(director = str_trim(director)) %>% 
  count(director, release_year, sort = TRUE)
```



Make a bar plot of the top directors with the most movies on netflix. Label axes and title the plot. Fill by the director. Get rid of the guide for the fill.
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  separate_rows(director, sep = ",") %>% 
  mutate(director = str_trim(director)) %>% 
  count(director, sort = TRUE) %>% 
  top_n(10) %>% 
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  theme_minimal() +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies")
```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*
#### Yes, the plot uses the default colors for the fill.

Make a bar plot of the top directors with the most movies on netflix. Label axes and title the plot. Fill by the director. Get rid of the guide for the fill. Use the fill scale to change the colors to a colorblind friendly palette.
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  separate_rows(director, sep = ",") %>% 
  mutate(director = str_trim(director)) %>% 
  count(director, sort = TRUE) %>% 
  top_n(10) %>% 
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Directors with the Most Movies on Netflix", x = "Director", y = "Number of Movies")

```

*How about for just 2020*

Make a bar plot of the top 10 directors with the most movies on netflix in 2020. Label axes and title the plot. Fill by the director. Get rid of the guide for the fill. Use the fill scale to change the colors to a colorblind friendly palette. Make sure the top 10 feature is working
```{r}
netflix_movies %>% 
  filter(!is.na(director)) %>% 
  separate_rows(director, sep = ",") %>% 
  mutate(director = str_trim(director)) %>% 
  filter(release_year == 2020) %>% 
  count(director, sort = TRUE) %>% 
  top_n(9) %>% 
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Directors with the Most Movies on Netflix in 2020", x = "Director", y = "Number of Movies")
```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*

## Make the top 10 directors barplot into a Shiny App where you can toggle by year
```{r}
# Define UI
ui <- fluidPage(
  titlePanel("Top 10 Directors by Year"),
  sidebarLayout(
    sidebarPanel(
      selectInput("year", "Select Year", choices = unique(netflix_movies$release_year))
    ),
    mainPanel(
      plotOutput("barplot")
    )
  )
)

# Define server logic
server <- function(input, output) {
  output$barplot <- renderPlot({
netflix_movies %>%
  filter(director != "NA") %>%
  filter(release_year == input$year) %>%
  separate_rows(director, sep = ",") %>%
  mutate(director = str_trim(director)) %>%
  count(director, sort = TRUE) %>%
  slice_max(n, n = 10) %>%
  slice(1:10)%>%
      ggplot(aes(x = reorder(director, n), y = n, fill = director)) +
      geom_col() +
      theme_minimal() +
      scale_fill_brewer(palette = "Set3") +
      guides(fill = FALSE) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      labs(title = paste("Top 10 Directors with the Most Movies on Netflix in", input$year),
           x = "Director",
           y = "Number of Movies")
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```


*Now use ChatGPT, Copilot or previous notes to fix the code*

## That's it! Let's take a break and then move on to working on group projects.
