---
title: "Lab 12 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating & Set-up
1. Please make sure that you sit next to your group members for the rest of the quarter.
mrsimon@ucdavis.edu (Conservatory System)
2. Please set-up your computer as normal.
3. Don't forget to record your attendance in iClicker.  

## Warm-up
For this warm-up, please use the `heartrate` data. Be sure to label your plots!
```{r}
heartrate <- read_csv("data/heartrate.csv")
```
```{r}
heartrate <- heartrate %>% 
  pivot_longer(-patient, #patient will not move, keep them where they are
               names_to = "drug", #make a new column called "drug", moving all current columns into new column
               values_to="heartrate" #values moved to a new column called "heartrate"
               )
```


1. Make a plot that compares the heartrate of patients for drug a.
```{r}
heartrate %>% 
  filter(drug == "a") %>% 
  ggplot(aes(x = patient, y = heartrate)) +
  geom_col()+
  labs(title = "Patients by Heartrate on Drug A",
       x = "Patient",
       y = "Heartrate (bpm)")
```


2. Make a plot that compares heartrate (as a range) for each drug.
```{r}
heartrate %>% 
  ggplot(aes(x = patient, y = heartrate))+
  geom_boxplot() +
    labs(title = "Patients Heartrate Ranges",
       x = "Patient",
       y = "Heartrate (bpm)")
```


## Group project  
We will meet in our groups today to compare the data sets that we found so far. You should try to pick a data set that you think will be most appropriate; i.e. does it have a mix of variables, does it have a story to tell? Once you have a data set, think about the kinds of questions you might ask. Remember, in the end you will need to deliver a 5-6 minute presentation of your work. If you need help, please ask!  