---
title: "Warmup_1-30"
author: "Elisabeth Sellinger"
date: "2024-01-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load tidyverse:
```{r}
library(tidyverse)
```

## Warm-up
1. Load the `bison.csv` data.
```{r}
bison <- read.csv("data/bison.csv")
```

2. What are the dimensions and structure of the data?
```{r}
dim(bison)
```

```{r}
str(bison)
```

3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
```{r}
bison_new <- bison %>% 
select("data_code", "animal_sex", "animal_weight", "animal_yob")
```

4. Pull out the animals born between 1980-1990.
```{r}
bison_new_1980 <- bison_new %>% 
filter(between(animal_yob, 1980, 1990))
```


5. How many male and female bison are represented between 1980-1990?
```{r}
table(bison_new_1980$animal_sex)
```

6. Between 1980-1990, were males or females larger on average?
```{r}
bison_new_1980_F <- bison_new_1980 %>% 
  filter(animal_sex == "F")
mean(bison_new_1980_F$animal_weight) #mean of females
```

```{r}
bison_new_1980_M <- bison_new_1980 %>% 
  filter(animal_sex == "M")
mean(bison_new_1980_M$animal_weight) #mean of males
```
### On average, males are larger.
